---
title: "Supplementary Material for 'Winter Cover Cropping Effects on Soil Water-Holding Capacity Vary by Site'"
author: "Nichols et al. 2021"
date: "6/15/2021"
header-includes: #allows you to add in your own Latex packages
- \usepackage{float} #use the 'float' package
- \floatplacement{figure}{H}  #make every figure with caption = h, this was the fix
output: 
  pdf_document:
    fig_cap: yes
    keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
library(tidyverse)
library(PFIweeds2020)
library(PFIswhc)
library(scales)
library(patchwork)
library(knitr)
library(kableExtra)
library(maps)
library(ggthemes)


library(purrr)

library(lme4)
library(lmerTest)
library(emmeans)

```

# Supplemental Figure S1. Map of sites
```{r sitemap, fig.cap="Map of site locations in Iowa"}

pfi_red <- "#9d3c22"
pfi_green <- "#b2bb1e"
pfi_blue <- "#036cb6"
pfi_orng <- "#e87d1e"
pfi_brn <- "#574319"

map_iowa <- as_tibble(map_data('state')) %>%
  filter(region == "iowa")

map_county <- as_tibble(map_data('county')) %>%
  filter(region == "iowa")

map_county3 <-
  map_county %>% filter(subregion %in% c("boone", "greene", "washington"))

set.seed(3)

ggplot() +
  geom_polygon(
    data = map_iowa,
    aes(x = long, y = lat, group = group),
    color = "gray20",
    fill = "white"
  ) +
  geom_polygon(
    data = map_county,
    aes(x = long, y = lat, group = group),
    color = "gray80",
    fill = NA
  ) +
  # geom_polygon(
  #   data = map_county3,
  #   aes(x = long, y = lat, group = group),
  #   color = pfi_blue,
  #   fill = "white",
  #   size = 3
  # ) +
  geom_jitter(
    data = pfi_siteinfo %>% rename("long" = "lon"),
    aes(x = long, y = lat),
    size = 4,
    fill = pfi_orng,
    width = 0.07,
    pch = 21
  ) +
  geom_text(aes(x = -95.05, y = 42.105), size = 7, label = "West", color = "black") +
  geom_text(aes(x = -93.5, y = 42.315), size = 7, label = "Central", color = "black") +
  geom_text(aes(x = -91.7, y = 41.05), size = 7, label = "East", color = "black") +
  theme(
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank()
  ) +
  coord_quickmap() +
  theme_map()
```

# Supplemental Figure S2. Soil texture results
```{r texture, fig.cap="Soil texture components varied by trial and cover crop treatment, with the cover cropped plots having significantly more sand bolded orange color, and significantly less clay at the West-grain and East-grain trials both, commercial fields"}

dat <- 
  sare_texture %>% 
  left_join(sare_plotkey) %>% 
  unite(site_name, sys_trt, col = "site_sys", sep = "-")

om <- 
  sare_om %>% 
  left_join(sare_plotkey) %>% 
  unite(site_name, sys_trt, col = "site_sys", sep = "-")


pfi_red <- "#9d3c22"
pfi_green <- "#b2bb1e"
pfi_blue <- "#036cb6"
pfi_orng <- "#e87d1e"
pfi_brn <- "#574319"

myorder <- c("clay", "silt", "sand")

# cover cover plots and no-cover separate----------------------------------------------------------

dat_pie_cc <- 
  dat %>% 
  select(site_sys, cc_trt, clay, silt, sand) %>% 
  pivot_longer(clay:sand) %>% 
  group_by(site_sys, cc_trt, name) %>% 
  summarise(value  = mean(value, na.rm = T)) %>% 
  ungroup() %>% 
  pivot_wider(names_from =  name, values_from = value)  %>% 
  mutate_if(is.numeric, round, 0) %>% 
  mutate(sand = 100 - clay - silt) %>% 
  pivot_longer(clay:silt) %>% 
  mutate(value = value/100)


#--use purrr to amek labels
dat_pie2_cc <- 
  dat_pie_cc %>% 
  mutate(name = factor(name, levels = myorder),
         name2 = fct_rev(name),
         value = round(value, 2)) %>% 
  arrange(name2) %>% 
  group_by(site_sys, cc_trt) %>% 
  nest() %>% 
  mutate(data2 = data %>% purrr::map(. %>% 
                                       arrange(name2) %>% 
                                       mutate(half = value/2,
                                              prev = lag(value),
                                              prev = ifelse(is.na(prev), 0, prev),
                                              cumprev = cumsum(prev),
                                              pos = half + cumprev) %>% 
                                       select(-c(half, prev, cumprev)))) %>% 
  unnest(data2)



#--fig
dat_pie2_cc %>% 
  # left_join(om %>% 
  #             group_by(site_sys, cc_trt) %>% 
  #             summarise(om = round(mean(om, na.rm = T), 1))) %>% 
  mutate(
    cc_trt = case_when(
      grepl("cc", cc_trt) ~ "Rye Cover Crop",
      grepl("no", cc_trt) ~ "No Cover",
      TRUE ~ cc_trt),
    site = site_sys,
    site_sys = factor(site_sys, levels = c("West-grain", "Central-silage", "Central-grain", "East-grain")) 
    #site = paste0(site, "\n", om, "% OM")
  ) %>% 
  ggplot(aes(x = "", y = value, fill = name, alpha = name)) +
  geom_bar(stat = "identity") +
  coord_polar("y", start = 0) +
  geom_text(aes(y = pos,
                label = paste(round(value, 4)*100, "%")),
                #label = percent(value, accuracy = 4)), 
            size = 5) +
  scale_fill_manual(values = c("clay" = pfi_brn, "silt" = pfi_red, "sand" = pfi_orng)) + 
  facet_grid(cc_trt~site_sys) +
  scale_alpha_manual(values = c(0.4, 0.4, 1)) +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.title = element_blank(),
        strip.text = element_text(size = rel(1.5)),
        legend.text = element_text(size = rel(1.3)),
        legend.position = "bottom") +
  labs(fill = NULL) +
  guides(alpha = F)
  

```

Table with values:
```{r texture2}

dat <- 
  sare_texture %>% 
  left_join(sare_plotkey) %>% 
  unite(site_name, sys_trt, col = "site_sys", sep = "-")

dat_pie_cc <- 
  dat %>% 
  select(site_sys, cc_trt, clay, silt, sand) %>% 
  pivot_longer(clay:sand) %>% 
  group_by(site_sys, cc_trt, name) %>% 
  summarise(value  = mean(value, na.rm = T)) %>% 
  ungroup() %>% 
  pivot_wider(names_from =  name, values_from = value)  %>% 
  mutate_if(is.numeric, round, 0) %>% 
  mutate(sand = 100 - clay - silt) %>% 
  pivot_longer(clay:silt) %>% 
  mutate(value = value/100)

dat_pie_cc %>% 
  pivot_wider(names_from = name, values_from = value) %>% 
  mutate(tot = clay + sand + silt,
         sig_sand_diff = ifelse(site_sys %in% c("East-grain", "West-grain"), "Yes", "No")) %>% 
  kable(caption = "Table of values", booktabs = T, align = "c") %>% 
  collapse_rows(columns = c(1,7), latex_hline = "major") %>% 
  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = F) 

```


# Supplemental Figure S3. Macro/micropore results
```{r pores, fig.cap="Mean percentage macropores (>30 µm) for each trial; percent macropores in the cover crop treatment was significantly lower (p=0.04) at the West-grain trial compared to the control treatment"}

pfi_red <- "#9d3c22"
pfi_green <- "#b2bb1e"
pfi_blue <- "#036cb6"
pfi_ltblue <- "#8DC4EB"
pfi_orng <- "#e87d1e"
pfi_brn <- "#574319"



# data --------------------------------------------------------------------

dp <- read_csv("../01_fit-models/dat_poresizes.csv") %>% 
  group_by(pore_cat, site_sys, cc_trt, rep_id) %>% 
  summarise(pct = sum(pct, na.rm = T))



# do macro/micro summary --------------------------------------------------
raws <- 
  dp %>% 
  filter(pore_cat == "Macropores (>30 um)") %>% 
  mutate(
    site_sys = factor(site_sys, 
                      levels = c("West-grain",
                                 "Central-silage", 
                                 "Central-grain",
                                 "East-grain")))


#--not sure which order to calculate them (lag/lead)

dat_pores <- 
  dp %>% 
  filter(pore_cat == 
           "Micropores (<30 um)") %>% 
  #--get mean for each trial
  group_by(pore_cat, site_sys, cc_trt) %>% 
  summarise(pct = mean(pct, na.rm = T)) %>% 
  pivot_wider(names_from = pore_cat, values_from = pct) %>% 
  #--force the macro to fill in the rest
  mutate(`Macropores (>30 um)` = 1 - `Micropores (<30 um)`) %>% 
    pivot_longer(3:4, names_to = "pore_cat", values_to = "pct") %>% 
  mutate(cc_trt = fct_rev(cc_trt)) %>% 
  mutate(
    site_sys = factor(site_sys, 
                      levels = c("West-grain",
                                 "Central-silage", 
                                 "Central-grain",
                                 "East-grain")),
    pore_cat = factor(pore_cat, 
                      levels = c("Macropores (>30 um)",
                                 "Micropores (<30 um)")),
    pore_cat = fct_rev(pore_cat))
  



# pie charts --------------------------------------------------------------

dat_pie <- 
  dat_pores %>% 
  mutate(pore_cat = fct_inorder(pore_cat),
         pore_cat = fct_rev(pore_cat)) %>% 
  group_by(site_sys, cc_trt, pore_cat) %>% 
  summarise(value = mean(pct, na.rm = T)) %>% 
  arrange(pore_cat) %>% 
  group_by(site_sys, cc_trt) %>% 
  nest() %>% 
  mutate(data2 = data %>% purrr::map(. %>%
                                       arrange(pore_cat) %>% 
                                       mutate(half = value/2,
                                              prev = lag(value),
                                              prev = ifelse(is.na(prev), 0, prev),
                                              cumprev = cumsum(prev),
                                              pos = half + cumprev) %>% 
                                       select(-c(half, prev, cumprev)))) %>% 
  unnest(data2)



# pie chart ---------------------------------------------------------------

dat_pie %>% 
  mutate(site = site_sys,
         pos2 = 1-pos,
         value2 = ifelse(pore_cat == "Micropores (<30 um)", NA, value),
         clr = paste(cc_trt, pore_cat)) %>% 
  ggplot(aes(x = "", y = value, fill = pore_cat)) +
  geom_bar(stat = "identity", color = "black") +
  coord_polar("y", start = 0) +
  geom_text(aes(y = pos2,
                label = percent(value2, accuracy = 2)), size = 5, show.legend = FALSE) +
  scale_fill_manual(values = c(pfi_ltblue, "white")) + 
  facet_grid(cc_trt~site) +
  guides(alpha = F) +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.title = element_blank(),
        strip.text = element_text(size = rel(1.1)),
        legend.text = element_text(size = rel(1.1)),
        legend.position = "bottom") +
  labs(fill = NULL)


```


# General Site Management Summary

```{r gentbl}

pfi_tb <- pfi_sitemgmt %>%
  rename(
    "Site Description" = site_desc, 
    "General Location" = loc,
    "Treatment Description" = trt_desc,
    "Year of Initiation" = initiation_year,
    "Crop Planted in 2019" = crop2019,
    "Number of Treatment Replicates" = reps_nu,
    "Sampled in 2019" = sampled2019,
    "Herbicides Used in 2018 Growing Season" = herb_growingseason2018,
    "Herbicdes Used in Fall 2018" = herb_fall2018,
    "Herbicides Used in Spring 2019" = herb_spring2019,
    "General Herbicide Regime" = herbspring_desc,
    "General Date of Cover Crop Termination" = herbspring2019_date,
    "General Date of Crop Planting" = cropplanting2019_date,
    "Inorganic Fertilizer Used" = inorganic_fert,
    "Organic Fertilizer Used" = organic_fert,
    "Tillage Used" = tillage
    ) 

#--general
pfi_tb %>% 
  select(1:7) %>% 
  kable(caption = "General Site Description", booktabs = T, align = "c") %>% 
  collapse_rows(columns = 1, latex_hline = "major") %>% 
  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = F) %>% 
  column_spec(column = c(1), width = "5em")  %>% 
  column_spec(column = c(2), width = "5em")  %>% 
  column_spec(column = c(3), width = "5em")  %>% 
  column_spec(column = c(4:6), width = "3em") 
```

\newpage

```{r herbtable}
#--herbicide
pfi_tb %>% 
  select(1, 8:10) %>% 
  kable(caption = "2018-2019 Herbicide Use", booktabs = T, align = "c") %>% 
  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = F) %>% 
  column_spec(column = c(1:4), width = "8em") %>% 
  collapse_rows(columns = 1, latex_hline = "major")
```

\newpage

```{r mgmttable}

#--mgmt
pfi_tb %>% 
  select(1, 11:16) %>% 
  kable(caption = "General Management", booktabs = T, align = "c") %>% 
  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = F) %>% 
  column_spec(column = c(2), width = "7em") %>% 
   column_spec(column = c(3:7), width = "5em") %>% 
   collapse_rows(columns = 1, latex_hline = "major")
  
```


\newpage
# Cover crop biomass production over past 10 years of trials

```{r ccbio}

  pfi_ccbio %>% 
  filter(year > 2009) %>% 
  mutate(site_name = recode(site_name,
                            "Boyd" = "Central",
                            "Funcke" = "West",
                            "Stout" = "East")) %>% 
  unite(site_name, sys_trt, col = "trial") %>% 
  group_by(trial, year) %>% 
  summarise(mbio = mean(ccbio_Mgha)) %>% 
  mutate_if(is.numeric, round, 2) %>% 
  pivot_wider(names_from = year, values_from = mbio) %>% 
  ungroup() %>% 
      kable(caption = "Historical cover crop biomass production (Mg/ha) by trial", booktabs = T, align = "c") %>% 
  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = F) %>% 
   collapse_rows(columns = 1, latex_hline = "major")

```



\newpage
# Statistical Results
 


```{r clay}
clay.stats <- read_csv("../01_fit-models/dat_clay-stats.csv")

clay.stats %>% 
  mutate_if(is.numeric, round, 3) %>% 
  mutate(diff_pval = ifelse(diff_pval < 0.001, "<0.001", diff_pval)) %>% 
  kable(caption = "Statistical analysis of cover crop effect on clay", booktabs = T, align = "c") %>% 
  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = F) 

sand.stats <- read_csv("../01_fit-models/dat_sand-stats.csv")

sand.stats %>% 
  mutate_if(is.numeric, round, 3) %>% 
  mutate(diff_pval = ifelse(diff_pval < 0.001, "<0.001", diff_pval)) %>% 
  kable(caption = "Statistical analysis of cover crop effect on sand", booktabs = T, align = "c") %>% 
  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = F) 


```

```{r omstats}

om.stats <- read_csv("../01_fit-models/dat_om-stats.csv")

om.stats %>% 
  mutate_if(is.numeric, round, 3) %>% 
  mutate(diff_pval = ifelse(diff_pval < 0.001, "<0.001", diff_pval)) %>% 
  kable(caption = "Statistical analysis of cover crop effect on organic matter, with and without sand covariate", booktabs = T, align = "c") %>% 
  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = F) %>% 
   collapse_rows(columns = c(1), latex_hline = "major")



```

```{r bulkden}

#--bd
PFIswhc::sare_bulkden %>% 
  left_join(sare_plotkey) %>% 
   group_by(site_name, sys_trt, crop_trt, cc_trt) %>% 
  summarise(bulkden_mean = mean(bulkden_gcm3, na.rm = T),
            bulkden_sd = sd(bulkden_gcm3, na.rm = T)) %>% 
  mutate_if(is.numeric, round, 2) %>% 
  kable(caption = "Mean bulk density (g/cm3) by trial", booktabs = T, align = "c") %>% 
  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = F) %>% 
   collapse_rows(columns = c(1), latex_hline = "major")

bd.stats <- read_csv("../01_fit-models/dat_bd-stats.csv")

bd.stats %>% 
  mutate_if(is.numeric, round, 3) %>% 
  mutate(diff_pval = ifelse(diff_pval < 0.001, "<0.001", diff_pval)) %>% 
  kable(caption = "Statistical analysis of cover crop effect on bulk density, with and without sand covariate", booktabs = T, align = "c") %>% 
  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = F) %>% 
   collapse_rows(columns = c(1), latex_hline = "major")



```

## Soil moisture (%vol) at saturation

```{r sat}


sat.stats <- read_csv("../01_fit-models/dat_sat-emmeans-diff.csv") %>% 
  select(cov, site_sys, everything(), -null.value)

sat.stats %>% 
  mutate_if(is.numeric, round, 3) %>% 
  mutate(adj.p.value = ifelse(adj.p.value < 0.001, "<0.001",adj.p.value)) %>% 
  kable(caption = "Statistical analysis of cover crop effect on soil water at saturation, with and without sand covariate", booktabs = T, align = "c") %>% 
  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = F) %>% 
   collapse_rows(columns = c(1), latex_hline = "major")
```
## Soil moisture (%vol) at field capacity (-100 cm water)
```{r fc}

fc.stats <- read_csv("../01_fit-models/dat_fc-emmeans-diff.csv") %>% 
  select(cov, site_sys, everything(), -null.value)

fc.stats %>% 
  mutate_if(is.numeric, round, 3) %>% 
  mutate(adj.p.value = ifelse(adj.p.value < 0.001, "<0.001",adj.p.value)) %>% 
  kable(caption = "Statistical analysis of cover crop effect on soil water at field capacity, with and without sand covariate", booktabs = T, align = "c") %>% 
  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = F) %>% 
   collapse_rows(columns = c(1), latex_hline = "major")
```
```{r paramsalp, fig.cap="Non-linear model fitted parameters"}

#--nonlinear fits
nlm.params <- 
  read_csv("../01_fit-models/dat_gard-parms-eu.csv") %>% 
  left_join(sare_plotkey) %>%
  unite(site_name, sys_trt, col = "site_sys") %>% 
  select(term, site_sys, cc_trt, rep, 
         estimate, std.error) %>% 
  arrange(term, site_sys, cc_trt, rep)

nlm.params %>% 
  ggplot(aes(rep, estimate, color = cc_trt)) + 
  geom_point() + 
  geom_linerange(aes(ymin = estimate - std.error, 
                     ymax = estimate + std.error)) +
  facet_grid(term ~ site_sys, scales = "free", labeller = label_wrap_gen())

```